---
title: "Github deliverable: The affects of temperature on the differeNT Memetes leaf colours"
author: "Duncan Sephton"
date: "2025-02-19"
output: html_document
---

# Packages that were installed to allow for the analysis.

install.packages("tidyverse"),
install.packages("ggplot2"),
install.packages("rmarkdown"),
install.packages("knitr"),


```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(rmarkdown)
library(knitr)
```

# Reading in the csv
```{r}
Jp <- read.csv("R_Deliverable.csv")
```
# Tidying up the data by changing the names in the column Colour
```{r}
Jp$colour[Jp$colour == "red - green"] <- "red_green"
Jp$colour[Jp$colour == "green-green"] <- "green_green"
```
## Ignoring all 0 values in the data set
```{r}
Jp <- Jp %>% filter(temp_bef != 0.0, temp_aft != 0.0, temp_change != 0.0)
```
# Removing unecessary data and making the boxplot
## Many columns were deamed unnecessary for the analysis, and due to the nature of the data, a boxplot was chosen to represent the data.
```{r}
Jp %>%
  group_by(colour) %>%
    select(!c("pair", "number", "shoot", "temp_ctr", "wet_weight", "time", "shoot_flower", "dry_weight")) %>% # removing the columns that are not needed
      pivot_longer(cols = c("temp_bef", "temp_aft", "temp_change"),
      names_to = "variable", values_to = "Temperature") %>% # making the data long
        ggplot() +     
        geom_boxplot(aes(y = Temperature, x = colour)) + # making the boxplot to have the temperature on the y axis and the colour on the x axis 
        ggtitle("Boxplot of temperature change by colour") +
        theme(plot.title = element_text(hjust=0.5))+
        labs(caption = "Figure 1: Boxplot showing the temperature (°C) before (temp_bef), after (temp_aft),
and total change in temperature (temp_change) of different Memetes leaf colours") +
        theme(plot.caption = element_text(hjust=0.5, face = "italic"))+
        theme(axis.text.x = element_text(angle = 45, hjust = 1))                                                       +      
        facet_wrap(vars(variable), scales = "free") # facet_wrap to have the boxplot for each variable         
```
